Embedding Dimension of Polyhedral Losses

### 2020-01-31 colt-20/embed-dim.tex
Adding another question on 1d things
Pass over abstract, intro, related work. bo@boulder

\begin{abstract}%
  A common technique in supervised learning with discrete losses, such as $0-1$ loss, is to optimize a convex surrogate loss over $\reals^d$, calibrated with respect to the original loss.
  In particular, recent work has investigated embedding the original predictions (e.g. labels) as points in $\reals^d$.
  %Recent work has proposed the notion of designing calibrated surrogate losses for classification-like problems through the lens of \emph{embedding} the original problem into $\reals^d$ and optimizing a polyhedral loss that is calibrated with respect to the original loss.
  In this work, we study the notion of \emph{embedding dimension} for given discrete losses.
  We characterize when a given discrete loss can be embedded into the real line, as well as when a higher-dimension input to the surrogate loss is required.
  Moreover, we give a quadratic feasibility program that yields lower bounds on the embedding dimension of a given discrete loss.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
In empirical risk minimization, the optimization of a loss function measuring error is canon in the design of algorithms designed for prediction tasks.
For many categorical tasks (e.g. classification, ranking, top-$k$, set inclusion), one can often use their intuition to design a discrete loss that models the question they want to answer about their data.
For example, $0-1$ loss $\ell(r)_y = \Ind{r \neq y}$ gives punishment $1$ for a misprediction and $0$ for a correct prediction.
This problem naturally lines up with predicting the canonical classification problem.
Since the optimization of discrete problems is computationally hard, we typically resort to optimizing a \emph{surrogate loss} that yields desirable statistical guarantees such as the existence of excess risk bounds given by \emph{consistency}.
In the finite outcome setting, we know that calibration implies consistency, so we study the design of calibrated, convex losses-- specifically polyhedral (piecewise linear and convex) losses.

\cite{finocchiaro2019embedding} propose the notion of \emph{embedding losses} in order to construct polyhedral, calibrated surrogate losses for a given discrete loss.
Their work constructively shows that every discrete loss is embeddable, but it is unclear how efficient an embedding can get.
In this work, we propose a notion of measuring the efficiency called \emph{embedding dimension} of this specific class of calibrated surrogate losses , which~\cite{finocchiaro2019embedding} show are well-equipped to be surrogates for discrete losses.
This notion of efficiency is measured by the dimension of the input to the surrogate, and in Section~\ref{sec:1d}, we characterize \jessie{and construct surrogates/links?} discrete losses that have a surrogate that takes input on the real line.
In Section~\ref{sec:d-dim}, we generalize our characterization to a general $d$-dimensional input for $d \geq 2$, presenting conditions on optimality and monotonicity that are necessary and sufficient for being able to construct a calibrated polyhedral surrogate taking $d$-dimensional input.
Moreover, we present a Quadratic Feasibility Program (Definition~\ref{def:qfp}) that has a solution if and only if the optimality condition of our general characterization is satisfied.
This feasibility program yields new techniques for finding lower bounds on this efficiency notion for a given discrete loss.
In Section~\ref{sec:examples}, we apply our characterizations to show new lower bounds on the embedding dimension for a problem whose convex calibration dimension (\cite{ramaswamy2016convex, ramaswamy2018consistent}) has been well-studied.
These results suggest that \emph{if} the convex calibration dimension of the same problems were lower than the embedding dimension, constructing a calibrated surrogate for such problems would require a new approach than previous works have taken.

\subsection{Related works}
\cite{bartlett2006convexity} study when a convex, calibrated surrogate exists for a given discrete loss, and in a finite-outcome setting,~\cite{tewari2007consistency} show that calibration (Definition~\ref{def:calibration}) is necessary and sufficient for consistency, meaning that there is a link mapping surrogate reports to discrete reports so that converging to the optimal prediction in the surrogate space implies the linked surrogate prediction converges to the optimal prediction for the original loss.

In order to decrease the dimension of the optimization problem,~\cite{frongillo2015elicitation} propose \emph{elicitation complexity}, which is the minimum dimension input to a loss such that, for all distributions $p$ over the outcomes $\Y$, the ``correct report,'' a function of the distribution $p$, minimizes the expected loss over the distribution $p$.
In the more specific case on constructing \emph{convex} calibrated surrogates for discrete losses,~\cite{ramaswamy2016convex} propose the notion of \emph{convex calibration dimension}, which is the minimum dimension input to a convex surrogate loss that is calibrated with respect to a given discrete loss.
While~\cite{ramaswamy2016convex} present methods for obtaining lower bounds on convex calibration dimension for any discrete loss, their method does not always work well, as demonstrated by the $\ellabs{\alpha}$ loss in Equation~\eqref{eq:abstain}.
\cite{ramaswamy2018consistent} present upper bounds on the convex calibration of $\abstain{\alpha}$ loss of $\ceil{\log_2(n)}$, but their characterization of lower bounds leave the tightness of these bounds as an open question.
These upper bounds are, however, much tighter than the previous upper bounds of $(n-1)$ given by the different surrogates presented by~\cite{crammer2001algorithmic} and~\cite{rifkin2004defense}, and it is unclear what characteristics of the abstain loss allow us to exponentially reduce dimension of the input.

In this work, we study \emph{embedding dimension}, which is the minimum dimension of input to a polyhedral surrogate that is calibrated with respect to the given discrete loss.
Moreover, we use this tool to characterize \emph{how} embeddable a given discrete loss is, showing necessary conditions for embedding a given loss into $d$ dimensions.
In order to concisely discuss embedding dimension, we appeal to notation and terminology from the field of property elicitation (\cite{frongillo2015vector-valued, gneiting2007strictly, lambert2009eliciting, osband1985information-eliciting, savage1971elicitation}), relating it to the language of calibrated surrogates as needed.

@@

\begin{theorem}
  Let $\ell: \R \to \nonnegreals^{\Y}$ be a discrete loss.
  Then $\ell$ is $1$-embeddable if and only if there is an ordering $\R = \{r_1,\ldots,r_k\}$ of the reports such that the following two conditions hold, where $v(i)_y := \ell(r_i)_y - \ell(r_{i-1})_y$:
  \begin{enumerate}
  \item For all $y\in\Y$, the sequence $\sgn(v(i)_y)$ is monotone in
    $i\in\{1,\ldots,k-1\}$,
  \item For all $i\in\{2,\ldots,k-1\}$, letting
    \begin{align*}
      \lambda^-(i) &= \min \left\{\frac{v(i)_y}{v({i+1})_y} : y\in\Y, v(i)_y, v({i+1})_y < 0\right\}
      \\
      \lambda^+(i) &= \max \left\{\frac{v(i)_y}{v({i+1})_y} : y\in\Y, v(i)_y, v({i+1})_y > 0\right\}~,
    \end{align*}
    we have $\lambda^-(i) \geq \lambda^+(i)$.
    (We adopt the convention $\max\emptyset = -\infty$, $\min\emptyset = +\infty$.)
  \end{enumerate}

>>>>

\begin{abstract}%
  A common technique in supervised learning with discrete losses, such as $0-1$ loss, is to optimize a convex surrogate loss over $\reals^d$, calibrated with respect to the original loss.
  In particular, recent work has investigated embedding the original predictions (e.g. labels) as points in $\reals^d$.
  %Recent work has proposed the notion of designing calibrated surrogate losses for classification-like problems through the lens of \emph{embedding} the original problem into $\reals^d$ and optimizing a polyhedral loss that is calibrated with respect to the original loss.
  In this work, we study the notion of \emph{embedding dimension} for given discrete losses.
  We characterize when a given discrete loss can be embedded into the real line, as well as when a higher-dimensional input to the surrogate loss is required.
  Moreover, we give a quadratic feasibility program that yields lower bounds on the embedding dimension of a given discrete loss.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
In supervised machine learning, the dominant paradigm is to measure performance of a model using some loss function $\ell($prediction, observation$)$.
In particular, this paper studies \emph{discrete} losses, in which the predictions lie in a finite set.
These are popular for many categorical tasks such as classification, ranking, top-$k$, set inclusion, and other structured prediction tasks.
%Models are trained using some variant of empirical risk minimization, i.e. minimization of loss on a training data set.
%For many categorical tasks (e.g. classification, ranking, top-$k$, set inclusion), one can often use intuition to design a discrete loss $\ell$ that models the question one wants to answer about one's data.
%For example, $0-1$ loss $\ell(r)_y = \Ind{r \neq y}$ gives punishment $1$ for a misprediction and $0$ for a correct prediction, and is used for classification problems over possibly large finite category sets.

However, the optimization of discrete losses over a dataset (i.e. empirical risk minimization) is typically computationally hard.
One therefore generally resorts to optimizing a computationally-nice \emph{surrogate loss} $L$.
The key requirement is \emph{calibration}, namely that optimizing surrogate loss allows one to recover the discrete prediction that optimizes the original $\ell$-loss.
Calibrated surrogate losses yields desirable statistical guarantees such as the existence of excess risk bounds given by consistency~\citep{tewari2007consistency}.
%In the finite outcome setting, we know that calibration implies consistency, so we study the design of calibrated, convex losses-- specifically polyhedral (piecewise linear and convex) losses.

The question therefore becomes: Given a discrete loss $\ell$, how do we design such calibrated surrogates $L$? 
The class of surrogates studied in this paper are \emph{polyhedral}, meaning piece-wise linear convex functions over prediction space $\reals^d$.
\cite{finocchiaro2019embedding} show that polyhedral surrogates are equivalent to the following recipe: \emph{embedding} the discrete loss $\ell$ into $\reals^d$ by letting each discrete prediction $r$ correspond to some point $\phi(r) \in \reals^d$, which optimizes $L$-loss if and only if $r$ optimizes $\ell$-loss.
\cite{finocchiaro2019embedding} show that every discrete loss $\ell$ can be embedded into \emph{some} dimension $d$.
But in the worst case $d=n-1$ where $n$ is the number of possible observations $y$, e.g. a number of classification categories.

%\emph{embedding losses} in order to construct polyhedral, calibrated surrogate losses for a given discrete loss.
%Their work constructively shows that every discrete loss is embeddable, but it is unclear how efficient an embedding can get.
In this work, we define and investigate the \emph{embedding dimension} of discrete losses.
A loss $\ell$ is $d$-embeddable if it can be embedded in $\reals^d$, equivalently, has some calibrated polyhedral surrogate $L$ over $\reals^d$.
Given $\ell$, it is desirable for computational and statistical reasons to construct a surrogate in as few dimensions as possible; it is also interesting to prove lower bounds, i.e. that no calibrated surrogate exists unless $d$ is large enough.

\paragraph{Contributions.}
In particular, $1$-embeddings have the simplest possible convex surrogate space, the real line.
In Section~\ref{sec:1d}, we characterize $1$-embeddability of a discrete loss $\ell$ via a variety of conceptual and testable conditions.
We give an efficient algorithm that, given the loss matrix of $\ell$, either proves that $\ell$ is not $1$-embeddable, or else constructs a calibrated $1$-dimensional polyhedral surrogate.
We also show that, perhaps surprisingly, for $d=1$, if \emph{any} convex calibrated surrogate exists, then in particular a polyhedral one does.
We conjecture that this continues to hold in higher dimensions.\bo{Formalize conjecture later?}
In other words, we conjecture that embedding dimension is actually always equal to the more permissive \emph{convex calibration dimension} of \cite{ramaswamy2016convex}.

In Section~\ref{sec:d-dim}, we observe a general characterization for $d \geq 2$, presenting conditions called optimality and monotonicity that are necessary and sufficient for $d$-embeddability.
A particular contribution is to isolate and investigate the optimality condition, which reduces to a finite set of conditions on the halfspace and vertex representations of a set of polytopes related to $\ell$ and $L$.
This allows us to construct a Quadratic Feasibility Program (Definition~\ref{def:qfp}) that has a solution if and only if the optimality condition can be satisfied.
In addition to providing a possible avenue toward automating the search for calibrated surrogate losses, this feasibility program yields new techniques for lower bounds on the embedding dimension of a given discrete loss.

In Section~\ref{sec:examples}, we apply our characterizations to show new lower bounds on the embedding dimension for a problem, classification with an abstain option, whose convex calibration dimension has been well-studied (\cite{ramaswamy2016convex, ramaswamy2018consistent}).
%These results suggest that \emph{if} the convex calibration dimension of the same problems were lower than the embedding dimension, constructing a calibrated surrogate for such problems would require a new approach than previous works have taken.
In particular we apply both our $1$-dimension characterization and higher-dimensional quadratic program to obtain previously-unknown lower bounds.

\subsection{Related work}
\cite{bartlett2006convexity} study when a convex, calibrated surrogate exists for a given discrete loss, and in a finite-outcome setting,~\cite{tewari2007consistency} show that calibration (Definition~\ref{def:calibration}) is necessary and sufficient for consistency.
%, meaning that there is a link mapping surrogate reports to discrete reports so that converging to the optimal prediction in the surrogate space implies the linked surrogate prediction converges to the optimal prediction for the original loss.

Several works have investigated the problem of reducing the dimensionality of a surrogate loss.
\cite{frongillo2015elicitation} propose \emph{elicitation complexity}, which roughly captures the minimum dimension $d$ of any surrogate loss $L$ (not necessarily convex or calibrated) for a given loss $\ell$ (not necessarily discrete).
In the more specific case of convex calibrated surrogates for discrete losses,~\cite{ramaswamy2016convex} propose the notion of \emph{convex calibration dimension}, which is the minimum dimension input to a convex surrogate loss that is calibrated with respect to a given discrete loss.
The difference in this paper is that our definition of embedding dimension requires the surrogate loss to additionally be polyhedral.
\cite{ramaswamy2016convex} present a condition for obtaining lower bounds on convex calibration dimension for any discrete loss, \emph{feasible subspace dimension}.
However, there are important examples, such as the $\ellabs{\alpha}$ loss in Equation~\eqref{eq:abstain}, for which feasible subspace dimension gives only trivial bounds.
A contribution of this work is to provide the first nontrivial lower bounds for this loss.

\cite{ramaswamy2018consistent} present upper bounds on the convex calibration of $\abstain{\alpha}$ loss of $\ceil{\log_2(n)}$, but their characterization of lower bounds leave the tightness of these bounds as an open question.
These upper bounds are, however, much tighter than the previous upper bounds of $(n-1)$ given by the different surrogates presented by~\cite{crammer2001algorithmic} and~\cite{rifkin2004defense}, and it is unclear what characteristics of the abstain loss allow us to exponentially reduce dimension of the input.

%In this work, we study \emph{embedding dimension}, which is the minimum dimension of input to a polyhedral surrogate that is calibrated with respect to the given discrete loss.
%Moreover, we use this tool to characterize \emph{how} embeddable a given discrete loss is, showing necessary conditions for embedding a given loss into $d$ dimensions.
In order to concisely discuss embedding dimension, we appeal to notation and terminology from the field of property elicitation (\cite{frongillo2015vector-valued, gneiting2007strictly, lambert2009eliciting, osband1985information-eliciting, savage1971elicitation}), relating it to the language of calibrated surrogates as needed.

@@

\begin{theorem}
  Let $\ell: \R \to \nonnegreals^{\Y}$ be a discrete loss.
  Then $\ell$ is $1$-embeddable if and only if there is an ordering $\R = \{r_1,\ldots,r_k\}$ of the reports such that the following two conditions hold, where $v(i)_y := \ell(r_i)_y - \ell(r_{i-1})_y$:
  \begin{enumerate}
  \item For all $y\in\Y$, the sequence $\sgn(v(i)_y)$ is monotone in
    $i\in\{1,\ldots,k-1\}$,
  \item For all $i\in\{2,\ldots,k-1\}$\jessie{why 2 to k-1?}, letting
    \begin{align*}
      \lambda^-(i) &= \min \left\{\frac{v(i)_y}{v({i+1})_y} : y\in\Y, v(i)_y, v({i+1})_y < 0\right\}
      \\
      \lambda^+(i) &= \max \left\{\frac{v(i)_y}{v({i+1})_y} : y\in\Y, v(i)_y, v({i+1})_y > 0\right\}~,
    \end{align*}
    we have $\lambda^-(i) \geq \lambda^+(i)$.
    (We adopt the convention $\max\emptyset = -\infty$, $\min\emptyset = +\infty$.)
  \end{enumerate}
